Airline Customer Satisfaction - Group 17

### 2025-04-04 Draft Business Analytics Assignment Group 17.Rmd
Draft Chapter 3

**dplyr** : used to filter rows, select columns, sort data, create new columns, and summarize data.

**tidyr** : imported to be used to clean and tidy the data, such as reshaping the data from wide to long format.

@@

**skimr** : designed to provide summary of statistical measures for each variables in data frames.

**DataExplorer** : used to display comprehensive and convenient interface for data exploration and data visualization.

@@

**ggplot2** : a package that is useful to visualize data in a more complex and customizable way compared to the basic R plotting functions.

**corrplot** : used to visualize correlation matrix between two variables in a more visually appealing and informative way.

@@

Below we will find the commands to install the required libraries so that the code can run properly.The libraries only need to be imported only once, later we only need to call the necessary libraries when needed on the specific sections.

@@

Recalling the Univariate Data Analysis, we will use five categories or variables that we deemed important as the base for further analysis, specifically to see the association between two categorical variables. These variables are distribution of gender, customer type, type of travel, class, and the disparity of numbers between satisfacted and unsatisfied passengers.

```{r}
plot_bar(data,ncol=2)
```
We will use the plot_bar function to visualize the association between these variables. For each of the following five sections, we take out one of the category and plot the association between each of the remaining four categories with the missing category. We start with creating the association between customer's satisfaction with the other four variables.

@@

From the univariate data distribution, we already know the numbers of satisfied and unsatisfied passengers is not equal with noticeable difference, with the majority of passengers being unsatisfied. So it is not surprising that regardless of the association between passengers' satisfaction with other categories, most graphs are showing the skew towards the neutral or dissatisfied category.

However, it could be noticed that the distribution of satisfaction level across both genders are relatively equal, while the first time flyers are more likely to be unsatisfied rather than the frequent travelers. Business travelers also tend to be more satisfied than the economy class passengers. This is also the case for the business class passengers. We will see in the other section that passengers with business intention are more likely to fly on business class, so the correlation is consistent in the satisfaction level as well. *This is understandable, since airlines with business class are more likely to have extra services and better quality of basic services, which means that it is more likely to satisfy the passengers. (belum tau info ini mending dimasukin di sini atau bagian data analysis).* Hence, the focus on improving the satisfaction rate of passengers should be on the economy class passengers with leisure intention, especially those who are first time flyers.

@@

In this section, we show the distribution of genders in four other categories, which is relatively balanced across all graphs. Hence, we can conclude that genders are not the deciding factor that we need to analyze further in the data analysis section, meaning that it is not important enough as a consideration point to improve the satisfaction rate.

@@

From the customer type distribution and association graphs, it can be noticed that the majority of first time passengers are flying economy class, and also more likely to have the less-favorable first imnpression towards their flying experience, which can be seen in its relation with satisfaction level.

@@

As we can see above, the majority of business travelers fly on the business class as well, and they are also more likely to be satisfied with their flying experience. This is consistent with our observation and discussion on the previous section. On the oppositem, travelers with personal reason, such as leisure, are more likely to fly on both of the economy classes and also less likely to be satisfied with their flying experience.

@@

**Rencana hapus section ini karena dirasa ga nambah observasi apapun, TBD**

@@

We noticed that the age distribution have high variation, therefore we try to create the age_group column to divide the age group into four categories: children, youth, adult, and senior. This categorization is based on most of the division range used in a lot of transportation services across Europe. The age group is defined and plotted in R as follows:

```{r}
data = data %>% mutate(Age_Range = if_else(Age <= 15,"Children",
if_else(16 <= Age & Age <= 25,"Youth",
if_else(26 < Age & Age <= 60,"Adult","Senior"))))
```

@@

Now we can interpret from the age group that the largest segment of our customer is belong to adult age group. This is understandable since the adult group has the widest range of age compared to other age groups.

Another data we can group is the flight distance which currently right skewed. We will create new variables of flight distance by grouping them into three categories: short-haul, medium-haul, and long-haul. The division of distance is based on the limit defined by David W. Wragg in his book A Dictionary of Aviation:

Short-haul Flights: 0 - 900 miles (0 - 1,450 km)

Medium-haul Flights: 900 - 2,200 miles (1,450 - 3,540 km)

Long-haul Flights: 2,200 miles (3,540 km) and above

```{r}
data = data %>% mutate(Distance_Group = if_else(Flight.Distance < 900,"Short-haul",
if_else(900 <= Flight.Distance & Flight.Distance <= 2200,"Medium-haul","Long-haul")))
```

plot_bar(data$Distance_Group, ncol = 2, order_bar = TRUE, ggtheme = theme_minimal())

We can divide which customer experiencing delay or not based on departure delay and arrival delay. United States' Federal Aviation Administration (FAA) defines a flight as delayed if it is 15 minutes or more behind schedule. We can use the following code to create the new variables to show the comparison of the numbers of flight with departure delay and arrival delay.

```{r}
data$departure.delay.status = ifelse(data$Departure.Delay > 15, "Delayed", "On Time")
data$arrival.delay.status = ifelse(data$Arrival.Delay > 15, "Delayed", "On Time")

plot_bar(data$departure.delay.status,ncol=2,order_bar=T,ggtheme = theme_minimal())
plot_bar(data$arrival.delay.status,ncol=2,order_bar=T,ggtheme = theme_minimal())
```

Because both departure and arrival delay status are equally important, we can combine them into one variable called **delay.status** with OR operator is used to display the total number of delayed flights.

```{r}
data$delay.status = ifelse(data$Departure.Delay > 15 | data$Arrival.Delay > 15, "Delayed", "On-time")

@@

Next, We group the features into category of services that could be improved by the airline:

**Flight General Service:** On-board Service, Food and Drink, In-flight Service

**Flight Entertainment:** In Flight Wifi Service, In Flight Entertainment

**Pre-flight Service:** Ease of Online Booking, Check-in Service, Online Boarding

**Comfortability:** Seat Comfort, Leg Room Service, Cleanliness

```{r}
data$In.flight.service = rowSums(data[,c('On.board.Service','Food.and.Drink','In.flight.Service')],na.rm = TRUE)
data$Flight.Entertainment = rowSums(data[,c('In.flight.Wifi.Service','In.flight.Entertainment')],na.rm = TRUE)
data$Pre.flight.service = rowSums(data[,c('Ease.of.Online.Booking','Check.in.Service','Online.Boarding')],na.rm = TRUE)
data$Comfortability = rowSums(data[,c('Seat.Comfort','Leg.Room.Service','Cleanliness')],na.rm = TRUE)

plot_histogram(data[ , c("In.flight.service", "Flight.Entertainment", "Pre.flight.service", "Comfortability")], ncol=2,ggtheme = theme_minimal())
```

Combining several category of services into one variable is useful to see the overall satisfaction of the airline services in numerous area of improvement. For example, as we can see above that in general the passengers found that the flights they are taking are relatively comfortable. However, the pre-flight service and flight entertainment category received mixed reviews and can be considered as an improvement area for the airline, where further analysis could be done to see which of the element of those services are the most dissatisfying for the passengers.

**DARI FEATURE ENGINEERING KE SECTION SELANJUTNYA HARUS CROSSCHECK, BEBERAPA VARIABEL BARU BERUBAH RANGE DAN PENAMAANNYA**

## Data Transformation <a name="2.4"></a>

@@

## Modeling preparation <a name="3.1"></a>

@@

@@

@@

@@

>>>>

**dplyr** : used to filter rows, select columns, sort data, create new columns, and summarize data.
**tidyr** : imported to be used to clean and tidy the data, such as reshaping the data from wide to long format.

@@

**skimr** : designed to provide summary of statistical measures for each variables in data frames.
**DataExplorer** : used to display comprehensive and convenient interface for data exploration and data visualization.

@@

**ggplot2** : a package that is useful to visualize data in a more complex and customizable way compared to the basic R plotting functions.
**corrplot** : used to visualize correlation matrix between two variables in a more visually appealing and informative way.

@@

Below we will find the commands to install and import the required libraries:

@@

```{r}
plot_bar(data,ncol=2)
```

Lorem ipsum

@@

Lorem ipsum

@@

Lorem ipsum

@@

Lorem ipsum

@@

Lorem ipsum

@@

Lorem ipsum

@@

We notice that the age distribution have high variation, therefore we try to create the age_group column

```{r}
data = data %>% mutate(Age_Range = if_else(Age <= 16,"Children",
if_else(17 <= Age & Age <= 30,"Youth",
if_else(30 < Age & Age <= 45,"Middle Aged","Old"))))
```

@@

Now we can interpret from the age group that the largest segment of our customer is belong to old age group which is >45 years old.

Another data we can group is the flight distance which currently right skewed

```{r}
data = data %>% mutate(Distance_Group = if_else(Flight.Distance < 1000,"Below 1000",
if_else(1000 <= Flight.Distance & Flight.Distance <= 2000,"1000-2000",
if_else(2000 < Flight.Distance & Flight.Distance < 3000,"2000-3000",
if_else(3000 < Flight.Distance & Flight.Distance < 4000,"3000-4000","Above 4000")))))
```

```{r}
plot_bar(data$Distance_Group,ncol=2,order_bar=T,ggtheme = theme_minimal())
```

We can divide which customer experiencing delay or not based on departure delay and arrival delay

```{r}
data$delay.status = ifelse(data$Departure.Delay > 0 | data$Arrival.Delay > 0, "Delayed", "On Time")

@@

We'll try to grouping the features into category of services that could be improved by the airline :

Flight General services : Food and Drink, In-flight Service
Flight Experience : In Flight Wifi Service, In Flight Entertainment
Digital services : Ease of Online Booking, Online Boarding
Comfort services : Seat Comfort, Leg Room Service, Cleanliness

```{r}
data$Flight.Service = rowSums(data[,c('Food.and.Drink','In.flight.Service')],na.rm = TRUE)
data$Flight.Entertainment = rowSums(data[,c('In.flight.Wifi.Service','In.flight.Entertainment')],na.rm = TRUE)
data$Digital.Service = rowSums(data[,c('Ease.of.Online.Booking','Online.Boarding')],na.rm = TRUE)
data$Comfort.Service = rowSums(data[,c('Seat.Comfort','Leg.Room.Service','Cleanliness')],na.rm = TRUE)

plot_histogram(data[ , c("Flight.Service", "Flight.Entertainment", "Digital.Service", "Comfort.Service")], ncol=2,ggtheme = theme_minimal())
```

## Data Tranformation <a name="2.4"></a>

@@

## Dimension Reduction <a name="3.1"></a>

PCA can be used to reduce the dimensionality of the dataset while ensuring that we dont lose too much information

**PCA with Numeric Variable Only**

```{r}
# Use the standardized data
data_standardized <- data %>%
  select(-c(ID, Arrival.Delay.Duration.log, Arrival.Delay.Duration.log.z,
            Departure.Delay.Duration.log, Departure.Delay.Duration.log.z, Flight.Distance.log,
            Flight.Distance.log.z,Departure.Delay.Duration, Arrival.Delay.Duration,
            Age.log, Age.log.z)) %>% 
  select(where(is.numeric)) %>%
  scale(center = TRUE, scale = TRUE)

results <- prcomp(data_standardized)
summary(results)
```

Cummulative value already achieved 90% using 12 Principal Components as below

```{r}
results$rotation[ , c(1:2) ]
```
To focus on variable with high loadings, we set the threshold of 0.2 and only select the first principal component.

```{r}
loadings <- data.frame(results$rotation)
threshold <- 0.20
indices <- abs(loadings$PC1) >= threshold
data.frame(variable = rownames(loadings)[indices], 
           loading = loadings$PC1[indices])
```

@@

```{r}
loadings_df <- data.frame(
  variable = rownames(loadings)[indices],
  loading = loadings$PC1[indices]
)

# Plot
ggplot(loadings_df, aes(x = reorder(variable, -loading), y = loading, fill = loading > 0)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  labs(title = "Top Loadings for PC1",
       x = "Variable",
       y = "Loading Value") +
  scale_fill_manual(values = c("steelblue", "tomato")) +
  theme_minimal()
```

Analyzing top contributor in PC1-PC12

```{r}
loadings <- as.data.frame(results$rotation)
loadings$Variable <- rownames(loadings)

# Top contributors for PC1–PC12
top_vars <- apply(abs(loadings[, 1:12]), 2, function(x) names(sort(x, decreasing = TRUE)[1:5]))
print(top_vars)
```
PC1 – In-Flight Amenities & Comfort
Driven by Flight.Entertainment, Comfort.Service, In.flight.Entertainment, Flight.Service, and Cleanliness. This component captures the overall in-flight experience focused on comfort and entertainment.

PC2 – Digital & Booking Experience
Influenced by Ease.of.Online.Booking, Digital.Service, and In.flight.Wifi.Service. It represents passengers’ interaction with digital services before and during the flight.

PC3 – Operational Onboard Service
Dominated by In.flight.Service, Baggage.Handling, and Seat.Comfort, this component reflects service reliability and seat-related comfort during the flight.

PC4 – Flight Timing & Boarding
Characterized by Arrival.Delay, Departure.Delay, and Online.Boarding. It relates to punctuality and boarding process efficiency.

PC5 – Passenger Profile & Journey
Shaped by Online.Boarding, Flight.Distance, Age, and Gate.Location, this component reflects demographic and journey-specific traits.

PC6 – Schedule & Gate Convenience
Led by Departure.and.Arrival.Time.Convenience, Gate.Location, and Age, this component captures how convenient flight times and gate accessibility are for passengers.

PC7 – Pre-Flight & Cabin Space
Defined by Check.in.Service, Leg.Room.Service, and Flight.Distance, combining service before boarding with cabin comfort.

PC8 – Age-Based Travel Patterns
Mainly driven by Age, Flight.Distance, and Check.in.Service, this component shows how age may influence travel experience and expectations.

PC9 – Comfort in Long Trips
Includes Leg.Room.Service, Flight.Distance, and Comfort.Service, highlighting comfort needs for longer or more frequent flights.

PC10 – In-Flight Consumables
Composed of On.board.Service, Flight.Service, and Food.and.Drink, focusing on passengers’ satisfaction with in-flight consumables.

PC11 – Delays & Terminal Logistics
Influenced by Arrival.Delay, Departure.Delay, and Gate.Location, this component represents disruptions tied to flight and gate logistics.

PC12 – Check-In & Time Coordination
Driven by Departure.and.Arrival.Time.Convenience and Check.in.Service, highlighting the efficiency of early-stage travel processes.

```{r}
screeplot(results, type = "lines")
```

**PCA with One Hot Encoding**

```{r}
library(caret)

# Use the standardized data
data_standardized <- data %>%
  select(-c(ID, Arrival.Delay.Duration.log, Arrival.Delay.Duration.log.z,
            Departure.Delay.Duration.log, Departure.Delay.Duration.log.z, Flight.Distance.log,
            Flight.Distance.log.z,Departure.Delay.Duration, Arrival.Delay.Duration,
            Age.log, Age.log.z))

dummies <- dummyVars(~ ., data = data_standardized)
data_encoded <- predict(dummies, newdata = data_standardized)

data_encoded_standardized <- scale(data_encoded, center = TRUE, scale = TRUE)

results <- prcomp(data_encoded_standardized)
summary(results)
```

When PCA was applied to numeric-only variables, the first 12 components explained over 90% of the variance, indicating strong inter-variable correlation and compact data structure.
However, after including one-hot encoded categorical variables, the number of components required to achieve the same level of variance increased to 20. This suggests that categorical variables contribute additional but more dispersed variance, requiring more dimensions to capture the overall structure.

@@

### Model Preparation <a name="3.2.1"></a>

```{r}
# Create copy of dataset for the purpose of classification only, convert to factor
data_clsf <- as.data.frame(lapply(data, as.factor))

# Encode Target Variable
data_clsf$Satisfaction <- ifelse(data_clsf$Satisfaction=="Satisfied",1,0)
data_clsf$Satisfaction <- factor(data_clsf$Satisfaction, levels = c(0, 1))

# For experiment, create subset of untransformed data, data with log, and data with normalization
data_clsf_ori <- data_clsf %>%
  select(-c(ID, Arrival.Delay.Duration.log, Arrival.Delay.Duration.log.z,
            Departure.Delay.Duration.log, Departure.Delay.Duration.log.z, Flight.Distance.log,
            Flight.Distance.log.z, Age.log, Age.log.z))

data_clsf_log <- data_clsf %>%
  select(-c(ID, Arrival.Delay.Duration.log.z, Departure.Delay.Duration.log.z, Flight.Distance.log.z,
            Departure.Delay.Duration, Arrival.Delay.Duration, Age.log.z, Arrival.Delay.Duration,
            Departure.Delay.Duration, Flight.Distance, Age))

data_clsf_norm <- data_clsf %>%
  select(-c(ID, Arrival.Delay.Duration.log,
            Departure.Delay.Duration.log, Flight.Distance.log,
            Age.log, Arrival.Delay.Duration,
            Departure.Delay.Duration, Flight.Distance, Age)) 
```

**Start with Untransformed Dataset**

**Split Dataset Training Testing**

```{r}
set.seed(46748717)

library (rsample)

proportion <-  0.7
split <- initial_split(data_clsf_ori, prop = proportion)
training <- training(split)
testing <- testing(split)
```

### Training a decision tree model on the data <a name="3.2.2"></a>

```{r}
model <- C50::C5.0(Satisfaction ~., 
                   data = training)
summary(model)
```

```{r}
print("Confusion matrix based on testing data")
pred.test <- predict(model, testing)
gmodels::CrossTable(testing$Satisfaction, pred.test,
                    prop.chisq = FALSE,
                    prop.c = FALSE,
                    prop.r = FALSE,
                    prop.t = FALSE,
                    dnn = c("Actual Satisfaction", "Predicted Satisfaction"))
```

```{r}
# Calculating Accuracy

accuracy = (21498 + 15676) / 38964
print(accuracy)
```

Accuracy already high with 95.4%

**Try with dataset that already log transformed**

```{r}
set.seed(46748717)

library (rsample)

proportion <-  0.7
split <- initial_split(data_clsf_log, prop = proportion)
training <- training(split)
testing <- testing(split)

model <- C50::C5.0(Satisfaction ~., 
                   data = training)
summary(model)

print("Confusion matrix based on testing data")
pred.test <- predict(model, testing)
gmodels::CrossTable(testing$Satisfaction, pred.test,
                    prop.chisq = FALSE,
                    prop.c = FALSE,
                    prop.r = FALSE,
                    prop.t = FALSE,
                    dnn = c("Actual Satisfaction", "Predicted Satisfaction"))
```

The result is almost identical since the algorithm determine the transformed data is not have significant importance. Therefore the improvement will focus on tuning the parameter of the model.

### Model Improvement <a name="3.3.3"></a>

**Boosting**

```{r}
library(C50)
library(gmodels)

set.seed(46748717)

proportion <-  0.7
split <- initial_split(data_clsf_ori, prop = proportion)
training <- training(split)
testing <- testing(split)

model_boost <- C5.0(Satisfaction ~.,
                    data = training,
                    trials = 10)

# Predicting on the testing data
print("Confusion matrix based on testing data (boosting)")
pred.test <- predict(model_boost, testing)
CrossTable(testing$Satisfaction, pred.test,
           prop.chisq = FALSE,
           prop.c = FALSE,
           prop.r = FALSE,
           prop.t = FALSE,
           dnn = c("Actual Satisfaction", "Predicted Satisfaction"))
```

```{r}
# Calculating Accuracy with Booster

accuracy = (21229 + 15816) / 38964
print(accuracy)
```

Accuracy didnt increase significant

**Assigning Cost to Mistake**

More costly to predict dissatisfied customer satisfied, cost higher in False Positive

```{r}
# Specifying the cost matrix
cost.matrix <- matrix(c(NA, 1,  # FN costs of predicting "Disatisfied" whereas actual value is "Satisfied"
                        2, NA), # FP costs of predicting "Satisfied" whereas actual value is "Disatisfeid"
                      nrow = 2, 
                      ncol = 2,
                      byrow = FALSE)
rownames(cost.matrix) <- colnames(cost.matrix) <- c(1, 0)

# Estimating the model with the cost matrix
model.cost <- C5.0(Satisfaction ~., 
              data = training,
              costs = cost.matrix)

print("Confusion matrix based on testing data (costs")
pred.test <- predict(model.cost, testing)
CrossTable(testing$Satisfaction, pred.test,
           prop.chisq = FALSE,
           prop.c = FALSE,
           prop.r = FALSE,
           prop.t = FALSE,
           dnn = c("Actual Satisfaction", "Predicted Satisfaction"))
```

```{r}
# Calculating Accuracy with Booster

accuracy = (21755  + 15087 ) / 38964
print(accuracy)
```

Accuracy decrease slightly

```{r}
# Calculating Decrease in False Positive

decrease_fp = (495   - 238  ) / 495
print(decrease_fp)
```

False positive decrease by 51.9%, from 495 to 238

**Random Forest**

```{r random forest}
data_randfor <- data_clsf_ori %>% select(-c(Arrival.Delay.Duration, Departure.Delay.Duration, Flight.Distance, Age))

proportion <-  0.7
split <- initial_split(data_randfor, prop = proportion)
training <- training(split)
testing <- testing(split)

model.forest <- randomForest::randomForest(Satisfaction ~., 
                                           data = training,
                                           ntree = 500, # how many trees should be grown?
                                           mtry = 2, # how many variables to sample at each split?
                                           replace = TRUE,
                                           importance = TRUE) # sampling of cases with or without replacement?

print("Confusion matrix based on testing data (costs")
pred.test <- predict(model.forest, testing)
CrossTable(testing$Satisfaction, pred.test,
           prop.chisq = FALSE,
           prop.c = FALSE,
           prop.r = FALSE,
           prop.t = FALSE,
           dnn = c("Actual Satisfaction", "Predicted Satisfaction"))
```

```{r}
# Calculating Accuracy with Random Forest

accuracy = (21410   + 15793  ) / 38964
print(accuracy)
```

Accuracy increase 95.4%

**Using Ranger to implement random forest to the database with high cardinality**

```{r}
install.packages("ranger")
library(ranger)
```

```{r}
proportion <-  0.7
split <- initial_split(data_clsf_ori, prop = proportion)
training <- training(split)
testing <- testing(split)

model.ranger <- ranger(Satisfaction ~., 
                       data = training,
                       num.trees = 500, # how many trees should be grown?
                       mtry = 2, # how many variables to sample at each split?
                       replace = TRUE) # sampling of cases with or without replacement?

print("Confusion matrix based on testing data (costs")
pred.test <- predict(model.ranger, testing)$predictions
CrossTable(testing$Satisfaction, pred.test,
           prop.chisq = FALSE,
           prop.c = FALSE,
           prop.r = FALSE,
           prop.t = FALSE,
           dnn = c("Actual Satisfaction", "Predicted Satisfaction"))
```

```{r}
# Calculating Accuracy with Ranger

accuracy = (21412   + 15739  ) / 38964
print(accuracy)
```

Accuracy didnt increase significant

### Variable Importance <a name="3.4"></a>

```{r}
library(randomForest)

importance <- importance(model.forest)
print(importance)
```

Plot the importance

```{r varimp-plot, fig.width=10, fig.height=8}
varImpPlot(model.forest)
```

High satisfaction on :
- Type of travel : Business
- Class : Business
- Customer Type : Returning

Type of Travel : Business

```{r fig.width=10, fig.height=8}
data_randfor <- data_clsf_ori[ which(data_clsf_ori$Type.of.Travel == "Business"), ]
data_randfor <- data_randfor %>% select(-c(Arrival.Delay.Duration, Departure.Delay.Duration,
                                           Flight.Distance, Age, Type.of.Travel))

proportion <-  0.7
split <- initial_split(data_randfor, prop = proportion)
training <- training(split)
testing <- testing(split)

model.forest <- randomForest::randomForest(Satisfaction ~., 
                                           data = training,
                                           ntree = 500, # how many trees should be grown?
                                           mtry = 2, # how many variables to sample at each split?
                                           replace = TRUE,
                                           importance = TRUE) # sampling of cases with or without replacement?

print("Confusion matrix for Random Forest Type of Travel Business")
pred.test <- predict(model.forest, testing)
CrossTable(testing$Satisfaction, pred.test,
           prop.chisq = FALSE,
           prop.c = FALSE,
           prop.r = FALSE,
           prop.t = FALSE,
           dnn = c("Actual Satisfaction", "Predicted Satisfaction"))

```{r fig.width=10, fig.height=8}
varImpPlot(model.forest)
```

Class : Business

```{r}
data_randfor <- data_clsf_ori[ which(data_clsf_ori$Class == "Business"), ]
data_randfor <- data_randfor %>% select(-c(Arrival.Delay.Duration, Departure.Delay.Duration,
                                           Flight.Distance, Age, Class))

proportion <-  0.7
split <- initial_split(data_randfor, prop = proportion)
training <- training(split)
testing <- testing(split)

model.forest <- randomForest::randomForest(Satisfaction ~., 
                                           data = training,
                                           ntree = 500, # how many trees should be grown?
                                           mtry = 2, # how many variables to sample at each split?
                                           replace = TRUE,
                                           importance = TRUE) # sampling of cases with or without replacement?

print("Confusion matrix for Random Forest Class Business")
pred.test <- predict(model.forest, testing)
CrossTable(testing$Satisfaction, pred.test,
           prop.chisq = FALSE,
           prop.c = FALSE,
           prop.r = FALSE,
           prop.t = FALSE,
           dnn = c("Actual Satisfaction", "Predicted Satisfaction"))
```

```{r fig.width=10, fig.height=8}
varImpPlot(model.forest)
```

Customer Type : Returning

```{r fig.width=10, fig.height=8}
data_randfor <- data_clsf_ori[ which(data_clsf_ori$Customer.Type == "Returning"), ]
data_randfor <- data_randfor %>% select(-c(Arrival.Delay.Duration, Departure.Delay.Duration,
                                           Flight.Distance, Age, Customer.Type))

proportion <-  0.7
split <- initial_split(data_randfor, prop = proportion)
training <- training(split)
testing <- testing(split)

model.forest <- randomForest::randomForest(Satisfaction ~., 
                                           data = training,
                                           ntree = 500, # how many trees should be grown?
                                           mtry = 2, # how many variables to sample at each split?
                                           replace = TRUE,
                                           importance = TRUE) # sampling of cases with or without replacement?

print("Confusion matrix for Random Forest Customer Type Returning")
pred.test <- predict(model.forest, testing)
CrossTable(testing$Satisfaction, pred.test,
           prop.chisq = FALSE,
           prop.c = FALSE,
           prop.r = FALSE,
           prop.t = FALSE,
           dnn = c("Actual Satisfaction", "Predicted Satisfaction"))
```

```{r fig.width=10, fig.height=8}
varImpPlot(model.forest)
```

Low Satisfaction on :
- Type of Travel Personal
- Class Economy and Economy Plus
- Customer Type First Time

Type of travel personal

```{r}
data_randfor <- data_clsf_ori[ which(data_clsf_ori$Type.of.Travel == "Personal"), ]
data_randfor <- data_randfor %>% select(-c(Arrival.Delay.Duration, Departure.Delay.Duration,
                                           Flight.Distance, Age, Type.of.Travel))

proportion <-  0.7
split <- initial_split(data_randfor, prop = proportion)
training <- training(split)
testing <- testing(split)

model.forest <- randomForest::randomForest(Satisfaction ~., 
                                           data = training,
                                           ntree = 500, # how many trees should be grown?
                                           mtry = 2, # how many variables to sample at each split?
                                           replace = TRUE,
                                           importance = TRUE) # sampling of cases with or without replacement?

print("Confusion matrix for Random Forest Type of Travel Personal")
pred.test <- predict(model.forest, testing)
CrossTable(testing$Satisfaction, pred.test,
           prop.chisq = FALSE,
           prop.c = FALSE,
           prop.r = FALSE,
           prop.t = FALSE,
           dnn = c("Actual Satisfaction", "Predicted Satisfaction"))
```

```{r fig.width=10, fig.height=8}
varImpPlot(model.forest)
```

Class Economy and Economy Plus

```{r}
data_randfor <- data_clsf_ori[ which(data_clsf_ori$Class %in% c("Economy", "Economy Plus")), ]
data_randfor <- data_randfor %>% select(-c(Arrival.Delay.Duration, Departure.Delay.Duration,
                                           Flight.Distance, Age, Class))

proportion <-  0.7
split <- initial_split(data_randfor, prop = proportion)
training <- training(split)
testing <- testing(split)

model.forest <- randomForest::randomForest(Satisfaction ~., 
                                           data = training,
                                           ntree = 500, # how many trees should be grown?
                                           mtry = 2, # how many variables to sample at each split?
                                           replace = TRUE,
                                           importance = TRUE) # sampling of cases with or without replacement?

print("Confusion matrix for Random Forest Type of Class Economy and Economy Plus")
pred.test <- predict(model.forest, testing)
CrossTable(testing$Satisfaction, pred.test,
           prop.chisq = FALSE,
           prop.c = FALSE,
           prop.r = FALSE,
           prop.t = FALSE,
           dnn = c("Actual Satisfaction", "Predicted Satisfaction"))
```

```{r fig.width=10, fig.height=8}
varImpPlot(model.forest)
```

Customer Type First Time

```{r}
data_randfor <- data_clsf_ori[ which(data_clsf_ori$Customer.Type %in% "First-time"), ]
data_randfor <- data_randfor %>% select(-c(Arrival.Delay.Duration, Departure.Delay.Duration,
                                           Flight.Distance, Age, Customer.Type))

proportion <-  0.7
split <- initial_split(data_randfor, prop = proportion)
training <- training(split)
testing <- testing(split)

model.forest <- randomForest::randomForest(Satisfaction ~., 
                                           data = training,
                                           ntree = 500, # how many trees should be grown?
                                           mtry = 2, # how many variables to sample at each split?
                                           replace = TRUE,
                                           importance = TRUE) # sampling of cases with or without replacement?

print("Confusion matrix for Random Forest Type of Customer Type First Time")
pred.test <- predict(model.forest, testing)
CrossTable(testing$Satisfaction, pred.test,
           prop.chisq = FALSE,
           prop.c = FALSE,
           prop.r = FALSE,
           prop.t = FALSE,
           dnn = c("Actual Satisfaction", "Predicted Satisfaction"))
```

```{r fig.width=10, fig.height=8}
varImpPlot(model.forest)
```

@@

Why Logistic Regression?

### Model Preparation <a name="3.3.1"></a>

```{r}
# Split training and testing data

set.seed(46748756)

library (rsample)

# Create copy of dataset for the purpose of classification only, convert to factor
data_logreg <- data %>%
  mutate(across(-c(ID, Age, Flight.Distance, Departure.Delay, Arrival.Delay, 
                   Departure.Delay.Duration,
                   Arrival.Delay.Duration, Arrival.Delay.Duration.log, Arrival.Delay.Duration.log.z,
                   Departure.Delay.Duration.log, Departure.Delay.Duration.log.z, Flight.Distance.log,
                   Flight.Distance.log.z, Age.log, Age.log.z), as.factor)) %>% subset(
                     select = -c(ID, Departure.Delay.Duration,
                   Arrival.Delay.Duration, Arrival.Delay.Duration.log, Arrival.Delay.Duration.log.z,
                   Departure.Delay.Duration.log, Departure.Delay.Duration.log.z)
                     )

# Cleaning column from zero score that only exist in small quantity (most of them only have 1 data, cant be splitted to training and testing)
cols_to_clean <- c("On.board.Service", "Check.in.Service", "Seat.Comfort",
                   "In.flight.Service", "Gate.Location", "Cleanliness", 
                   "In.flight.Entertainment")

for (col in cols_to_clean) {
  data_logreg <- subset(data_logreg, data_logreg[[col]] != "0")
  data_logreg[[col]] <- droplevels(data_logreg[[col]])
}

proportion <-  0.7
split <- initial_split(data_logreg, prop = proportion)
training <- training(split)
testing <- testing(split)
```

@@

### Training Logistic Regression Model <a name="3.3.2"></a>

```{r}
# Using all data
model_log <- glm(Satisfaction ~., 
               data = training, family = "binomial")

summary(model_log)
```

### Interpretation of the Model <a name="3.3.3"></a>

**Interpret the coefficient**

```{r}
exp(coef(model_log))
```

**Visualize the coefficient**

```{r fig.width=10, fig.height=20}
coefs <- summary(model_log)$coefficients
coef_df <- data.frame(
  Variable = rownames(coefs),
  Estimate = coefs[, "Estimate"],
  StdError = coefs[, "Std. Error"],
  p_value = coefs[, "Pr(>|z|)"]
)

# Remove intercept for clarity
coef_df <- coef_df[coef_df$Variable != "(Intercept)", ]

# Plot using ggplot2
library(ggplot2)

ggplot(coef_df, aes(x = reorder(Variable, Estimate), y = Estimate)) +
  geom_point(color = "darkblue") +
  geom_errorbar(aes(ymin = Estimate - 1.96 * StdError,
                    ymax = Estimate + 1.96 * StdError), width = 0.2) +
  coord_flip() +
  labs(title = "Logistic Regression Coefficients",
       x = "Predictor",
       y = "Coefficient Estimate (log-odds)") +
  theme_minimal()
```

**Interpreting Z Value**

```{r}
model_summary <- summary(model_log)

# Extract z-values and convert to data frame
z_df <- data.frame(
  Variable = rownames(model_summary$coefficients),
  Z_value = model_summary$coefficients[, "z value"]
)

# Remove intercept
z_df <- subset(z_df, Variable != "(Intercept)")

z_df
```

```{r  fig.width=10, fig.height=20}
# Plot using ggplot2
library(ggplot2)

ggplot(z_df, aes(x = reorder(Variable, abs(Z_value)), y = Z_value)) +
  geom_col(fill = "steelblue") +
  geom_hline(yintercept = c(-1.96, 1.96), linetype = "dashed", color = "red") + # 95% significance threshold
  coord_flip() +
  labs(title = "Z-values from Logistic Regression",
       x = "Predictor Variable",
       y = "Z-value (Wald Statistic)") +
  theme_minimal()
```

### Model Evaluation <a name="3.3.4"></a>

```{r}
pred.test <- predict(model_log, newdata = testing, type = "response")

# Convert to class labels
pred.test.result <- ifelse(pred.test > 0.5, "Satisfied", "Neutral or Dissatisfied")
testing$Satisfaction <- factor(testing$Satisfaction, levels = c("Satisfied", "Neutral or Dissatisfied"))
pred.test.result <- factor(pred.test.result, levels = c("Satisfied", "Neutral or Dissatisfied"))

confusionMatrix(pred.test.result, testing$Satisfaction)
```

Accuracy 93.7% with sensitivity 95.2% and specificy 91.9%

**Interpret ROC and AUC**

```{r}
library(pROC)

# ROC curve and AUC
roc_obj <- roc(testing$Satisfaction, pred.test)
plot(roc_obj, col = "blue", main = "ROC Curve")
auc(roc_obj)
```

```{r}
auc(roc_obj)
```

```{r}
plot(model_log)
```

### Improving Model Performance

**Using significant variables only**

p-value < 0.05

```{r}
training_signif <- training %>% subset(select = -c(Departure.Delay, Arrival.Delay, Flight.Distance, 
                                                  In.flight.Wifi.Service, Food.and.Drink, 
                                                  In.flight.Entertainment, Age_Range, Distance_Group,
                                                  Flight.Service, Flight.Entertainment, 
                                                  Comfort.Service, Digital.Service, 
                                                  Flight.Distance.log.z, Age.log.z))

model_log_signif <- glm(Satisfaction ~., 
               data = training_signif, family = "binomial")

summary(model_log_signif)

AIC increase to 44749

```{r}
pred.test <- predict(model_log_signif, newdata = testing, type = "response")

# Convert to class labels
pred.test.result <- ifelse(pred.test > 0.5, "Satisfied", "Neutral or Dissatisfied")
testing$Satisfaction <- factor(testing$Satisfaction, levels = c("Satisfied", "Neutral or Dissatisfied"))
pred.test.result <- factor(pred.test.result, levels = c("Satisfied", "Neutral or Dissatisfied"))

confusionMatrix(pred.test.result, testing$Satisfaction)
```

The accuracy decrease

ANOVA test :

```{r}
anova(model_log, model_log_signif, test="Chisq")
```

Model full siginificantly better than model with significant only

**Adjusting Threshold**

```{r}
coords(roc_obj, "best", ret = "threshold")
```

```{r}
pred.test <- predict(model_log, newdata = testing, type = "response")
